canaccessR: An open data product for analyzing transportation accessibility to employment and grocery stores in Canada's largest metropolitan areas.

### 2025-03-18 paper/Manuscript-Data-Package.Rmd
AS edits to the draft

The objective of this paper is to describe the {canaccesR} open data package.
Its main contents are a set of public transit travel time matrices (TTM) estimates to employment and groceries stores from the 12 largest Canadian metropolitan areas in 2019 and 2023.
The results are stored at the Dissemination Areas (DA) [^1] level, yielding origin-destination pairs containing information on travel time by public transit, population, and total employment within each metropolitan area.
These estimates were created by leveraging expertise in data science, computer programming, and transportation accessibility using the {r5r} package [@pereiraR5rRapidRealistic2021].
We used public transit schedule, transport network, population, and business location data from different sources to estimate these TTM's datasets.

@@

Estimating accessibility - the potential offered by the transportation system to reach destinations [@paezMeasuringAccessibilityPositive2012] - requires specialized datasets and technical expertise.
Recent efforts following an open-source and transparent philosophy have been made to disseminate useful data and information on transportation in the Canadian context [@soukhovTTS2016RDataSet2023]. 
However, despite these initiatives, pre-processed and available data that allow for the ease estimation of accessibility indicators are still scarce.
Within this context, we expect to help filling this gap by the processing raw into user-ready data and making them publicly available to advance knowledge on the field.
Our main contribution is to provide analysis-ready data for Canada's largest cities on the topic of transportation accessibility, thus making urban analytics in the country more accessible and contributing to future research and data-based decision making.

@@

The locations included in the data package comprise the 12 largest (population-wise) Census metropolitan areas (CMA's) based on the 2016 Canadian Census [@governmentofcanada2016CensusPopulation2016] [^2]. 
These locations are Toronto, Montreal, Vancouver, Ottawa-Gatineau, Calgary, Edmonton, Quebec City, Winnipeg, Hamilton, Kitchener-Cambridge-Waterloo, London, and Halifax.
We used four main data sources to construct the {canaccesR} data package: General Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI’s Enhanced Points of Interest, and Statistics Canada Census data.

[^2]: We included Oshawa, Ontario, as part of the Greater Toronto Area (GTA) because of the former's proximity to the latter. We also included Abbotsford-Mission, British Columbia, as part of the Vancouver metropolitan area because of the former's proximity to a transit station on the region’s West Coast Express commuter rail line.

@@

Using the {r5r} package, we estimated public transit travel times for two destination types, grocery stores and jobs. 
For each amenity type, we chose a likely travel time and day of the week. 
We set a 15 minutes time window and the maximum trip duration to 120 minutes. 
The estimated times are the median of the 15 minute time window.
For groceries stores, we set the departure date to a weekend afternoon and the departure time to between 12:00 PM to 12:15 PM on April 20, 2019 and April 22, 2023.
For employment, we ran the analysis on a typical weekday morning rush-hour commute, more specifically 8:00 to 8:15 AM departure on Tuesday, April 16, 2019 and Tuesday, April 18, 2023 [^3].
In both cases, we assumed that walking was the mode of travel from origin to transit stop and from transit stop to destination.
We aggregated all the resulting travel time matrices at the Dissemination Area (DA) level, which comprise the fundamental unit of analysis in data package.

@@

<!-- ## Travel time matrices -->

The package contains the following contents: travel time matrices, socio-economic and demographic data disaggregated at the DA level, the CMA areas' boundaries and backgrounds for plotting the data spatially, and aggregated statistics.

The main contents of the {canaccessR} package are the travel time matrices.
<!-- These matrices are data frames with origin-destination pairs, travel time by public transit, and population and employment by Census Dissemination Area.  -->
These matrices (datasets) comprise the estimated travel time by transit from and to each origin and destination Dissemination Area pairs.
The datasets also feature the total population in the origin DA and the total employment in the destination DA.
Besides these information, the matrices contain the unique origin and destination DA codes, region and name identifiers, the year in which the travel times refers to, and the date and time of departure for a trip that originates at the Dissemination Area of origin.
There are two matrices for each metropolitan area, one containing the travel times to jobs and the other to groceries stores [^4].
These information can be turned into spatial data using the socio-economic and demographic datasets.

[^4]: The information from Toronto, Hamilton, and Waterloo are aggregated at the Greater Golden Horseshoe travel time matrices, making it 20 datasets in total.

The socio-economic and demographic data contain other information disaggregated at the DA level for each CMA. 
These refer to total population by age groups, number of dwellings, number of individuals below the Low Income Measure, etc.
The aggregated statistics refer to population aggregates (for the selected CMA's) and transit revenue and ridership data aggregated by regional and national scale.

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. 
The Table @tbl-table_1 below summarizes the travel time estimates.
The travel time datasets contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DA's in the sample and 18,519,897 pairs from population to groceries stores.
Considering all areas combined, the mean travel time to jobs was 80 and 76 for groceries stores.

@@

```{r, fig.pos="H"}
#| label: fig-travel_time_emp_grc_plot
#| fig-cap: "Estimated median travel time (minutes) per Dissemination Area to jobs (left) and groceries stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021). DA and planning boundaries of the Montréal metropolitan area (Statistics Canada, 2016)."
#| out-width: 100%
#| echo: false
knitr::include_graphics("../figures/patch_tt_emp_grc.jpg")
```

>>>>

The objective of this paper is to describe the {canaccesR} open data package. Its main contents are a set of public transit travel time matrices (TTM) estimates to employment and groceries stores from the 12 largest Canadian metropolitan areas in 2019 and 2023, representing approximately 55% of the Canadian population. The results are provided at the Dissemination Areas (DA) [^1] level, yielding origin-destination pairs containing associated travel time by public transit, population, total employment, mode share, and other relevant census variables and spatial shape boundaries for each metropolitan area. This data package was created by leveraging expertise in data science, R programming, and transportation analysis. It includes .Rmd notebooks written for computing travel time matrices for large sets of origin-destination pairs using the {r5r} R package [@pereiraR5rRapidRealistic2021]. Overall, {canaccesR} offers `r data(package = "canaccessR")$results[, "Item"] |> length()` complementary objects ready for temporal and spatial analysis. The package is an analysis-read product based on a fusion of data sources, including public transit schedules (GTFS files), road and transit networks, census data, and filtered business location data.

@@

TTMs are a core piece of information required for estimating spatial interaction. Accessibility -the _potential_ spatial interaction offered by the transportation system to reach destinations [@paezMeasuringAccessibilityPositive2012]- is an example of a measure that requires TTMs. Recent efforts following an open-source and transparent philosophy have been made to disseminate useful data and information on transportation in the Canadian context [@soukhovTTS2016RDataSet2023]. However, despite these initiatives, pre-processed and available data that allow for the ease estimation of accessibility indicators are still scarce. Within this context, we expect to help filling this gap by the processing raw into user-ready data and making them publicly available to advance knowledge on the field. Our main contribution is to provide analysis-ready data for Canada's largest cities on the topic of transportation accessibility, thus making urban analytics in the country more accessible and contributing to future research and data-based decision making.

@@

The locations included in the data package comprise the 12 largest (population-wise) Census metropolitan areas (CMA's) based on the 2016 Canadian Census [@governmentofcanada2016CensusPopulation2016] [^2].These locations are the surrunding Toronto, Montreal, Vancouver, Ottawa-Gatineau, Calgary, Edmonton, Quebec City, Winnipeg, Hamilton, Kitchener-Cambridge-Waterloo, London, and Halifax areas. We used four main data sources to construct the {canaccesR} data package: General Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI’s Enhanced Points of Interest, and Statistics Canada Census data.

[^2]: We included Oshawa as part of the Greater Toronto Area (GTA) because of its proximity. Similarly, we included Abbotsford-Mission as part of the Vancouver metropolitan area due to its proximity to a transit station on the region’s West Coast Express commuter rail line.

@@

Using the {r5r} R package, we estimated public transit travel times for two destination types, grocery stores and jobs.  For each amenity type, we chose a likely travel time and day of the week. We set a 15 minutes time window and the maximum trip duration to 120 minutes. The estimated times are the median of the 15 minute time window. For groceries stores, we set the departure date to a weekend afternoon and the departure time to between 12:00 PM to 12:15 PM on April 20, 2019 and April 22, 2023. For employment, we ran the analysis on a typical weekday morning rush-hour commute, more specifically 8:00 to 8:15 AM departure on Tuesday, April 16, 2019 and Tuesday, April 18, 2023 [^3]. In both cases, we assumed that walking was the mode of travel from origin to transit stop and from transit stop to destination. We aggregated all the resulting travel time matrices at the Dissemination Area (DA) level, which comprise the fundamental unit of analysis in data package.

@@

<!-- ## Travel time matrices -->
```{r}
data(package = "canaccessR")$results[, "Item"]
grep(paste("^", "census_data_da_", sep=""), data(package = "canaccessR")$results[, "Item"])
 data("census_data_cma_cal")
```

Specifically, the package contains the following contents: `r grep(paste("^", "travel_matrix_grc_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` data.frame objects containing the calculated public travel times from DA centroids to grocery stores and another `r grep(paste("^", "travel_matrix_emp_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` for DA centroids to DA centroids. Notably, the {travel_matrix_grc_ggh} data.frame, is named after the acronym for the Greater Golden Horseshoe area, which includes the CMA regions of Toronto, Hamilton, and Kitchener-Cambridge-Waterloo. Hence, the `r grep(paste("^", "travel_matrix_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` travel time data.frames represent public transit travel times for all 12 CMAs across two sets of destinations. Next, `r grep(paste("^", "census_data_da_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` sf objects represent census data for each DA, including dwelling counts, population by age bracket, single-parent-headed households, low-income prevalence, official language knowledge, housing- quality, ownership and affordability variables, visible minority, newcomer- and immigration- related variables, educational attainment, and commuting mode shares. Furthermore, `r grep(paste("^", "census_data_cma_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` and `r grep(paste("^", "region_background_", sep=""), data(package = "canaccessR")$results[, "Item"]) |> length()` sf objects represent the CMA areas' boundaries and backgrounds for plotting the data spatially, respectively. Finally, `r grep("statistics", data(package = "canaccessR")$results[, "Item"]) |> length()` data.frames contain aggregated population and transit statistics at the CMA level.

<!--AS: Joao, I rephrased the below into one paragraph, what do you think?  -->

<!-- The main contents of the {canaccessR} package are the travel time matrices. -->
<!-- <!-- These matrices are data frames with origin-destination pairs, travel time by public transit, and population and employment by Census Dissemination Area.  --> -->
<!-- These matrices (datasets) comprise the estimated travel time by transit from and to each origin and destination Dissemination Area pairs. The datasets also feature the total population in the origin DA and the total employment in the destination DA. Besides these information, the matrices contain the unique origin and destination DA codes, region and name identifiers, the year in which the travel times refers to, and the date and time of departure for a trip that originates at the Dissemination Area of origin. There are two matrices for each metropolitan area, one containing the travel times to jobs and the other to groceries stores [^4]. These information can be turned into spatial data using the socio-economic and demographic datasets. -->

<!-- [^4]: The information from Toronto, Hamilton, and Waterloo are aggregated at the Greater Golden Horseshoe travel time matrices, making it 20 datasets in total.  -->

<!-- The socio-economic and demographic data contain other information disaggregated at the DA level for each CMA.  -->
<!-- These refer to total population by age groups, number of dwellings, number of individuals below the Low Income Measure, etc. -->
<!-- The aggregated statistics refer to population aggregates (for the selected CMA's) and transit revenue and ridership data aggregated by regional and national scale. -->

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. Table @tbl-table_1 summarizes the travel time estimates. The travel time objects contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DA's in the sample and 18,519,897 pairs from population to groceries stores. Considering all areas combined, the mean travel time to jobs was 80 and 76 for groceries stores.

@@

```{r, fig.pos="H"}
#| label: fig-travel_time_emp_grc_plot
#| fig-cap: "Estimated median travel time (minutes) per Dissemination Area to jobs (left) and groceries stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021). DA and planning boundaries of the Montréal metropolitan area (Statistics Canada, 2016)."
#| out-width: 100%
#| echo: false
knitr::include_graphics("../figures/patch_tt_emp_grc.jpg")
```
<!--AS: Where is the 0-30 bin for grocery stores? It would be helpful to change the plot so it reads '0-30' '30-60', '60-90', 90-120min -->

### 2025-03-18 paper/Manuscript-Data-Package.tex
AS edits to the draft

The objective of this paper is to describe the \{canaccesR\} open data
package. Its main contents are a set of public transit travel time
matrices (TTM) estimates to employment and groceries stores from the 12
largest Canadian metropolitan areas in 2019 and 2023. The results are
stored at the Dissemination Areas (DA) \footnote{Dissemination Areas are
  the smallest publicly available spatial unit provided by Statistics
  Canada \citep{governmentofcanadaDictionaryCensusPopulation2021a}.}
level, yielding origin-destination pairs containing information on
travel time by public transit, population, and total employment within
each metropolitan area. These estimates were created by leveraging
expertise in data science, computer programming, and transportation
accessibility using the \{r5r\} package
\citep{pereiraR5rRapidRealistic2021}. We used public transit schedule,
transport network, population, and business location data from different
sources to estimate these TTM's datasets.

Estimating accessibility - the potential offered by the transportation
system to reach destinations
\citep{paezMeasuringAccessibilityPositive2012} - requires specialized
datasets and technical expertise. Recent efforts following an
open-source and transparent philosophy have been made to disseminate
useful data and information on transportation in the Canadian context
\citep{soukhovTTS2016RDataSet2023}. However, despite these initiatives,
pre-processed and available data that allow for the ease estimation of
accessibility indicators are still scarce. Within this context, we
expect to help filling this gap by the processing raw into user-ready
data and making them publicly available to advance knowledge on the
field. Our main contribution is to provide analysis-ready data for
Canada's largest cities on the topic of transportation accessibility,
thus making urban analytics in the country more accessible and
contributing to future research and data-based decision making.

@@

The locations included in the data package comprise the 12 largest
(population-wise) Census metropolitan areas (CMA's) based on the 2016
Canadian Census \citep{governmentofcanada2016CensusPopulation2016}
\footnote{We included Oshawa, Ontario, as part of the Greater Toronto
  Area (GTA) because of the former's proximity to the latter. We also
  included Abbotsford-Mission, British Columbia, as part of the
  Vancouver metropolitan area because of the former's proximity to a
  transit station on the region's West Coast Express commuter rail line.}.
These locations are Toronto, Montreal, Vancouver, Ottawa-Gatineau,
Calgary, Edmonton, Quebec City, Winnipeg, Hamilton,
Kitchener-Cambridge-Waterloo, London, and Halifax. We used four main
data sources to construct the \{canaccesR\} data package: General
Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI's Enhanced
Points of Interest, and Statistics Canada Census data.

@@

Using the \{r5r\} package, we estimated public transit travel times for
two destination types, grocery stores and jobs. For each amenity type,
we chose a likely travel time and day of the week. We set a 15 minutes
time window and the maximum trip duration to 120 minutes. The estimated
times are the median of the 15 minute time window. For groceries stores,
we set the departure date to a weekend afternoon and the departure time
to between 12:00 PM to 12:15 PM on April 20, 2019 and April 22, 2023.
For employment, we ran the analysis on a typical weekday morning
rush-hour commute, more specifically 8:00 to 8:15 AM departure on
Tuesday, April 16, 2019 and Tuesday, April 18, 2023 \footnote{The one
  exception is Quebec City, where the routing for 2019 occurs on a
  Saturday and Tuesday in June (instead of April) due to the GTFS data
  unavailability.}. In both cases, we assumed that walking was the mode
of travel from origin to transit stop and from transit stop to
destination. We aggregated all the resulting travel time matrices at the
Dissemination Area (DA) level, which comprise the fundamental unit of
analysis in data package.

@@

The package contains the following contents: travel time matrices,
socio-economic and demographic data disaggregated at the DA level, the
CMA areas' boundaries and backgrounds for plotting the data spatially,
and aggregated statistics.

The main contents of the \{canaccessR\} package are the travel time
matrices. These matrices (datasets) comprise the estimated travel time
by transit from and to each origin and destination Dissemination Area
pairs. The datasets also feature the total population in the origin DA
and the total employment in the destination DA. Besides these
information, the matrices contain the unique origin and destination DA
codes, region and name identifiers, the year in which the travel times
refers to, and the date and time of departure for a trip that originates
at the Dissemination Area of origin. There are two matrices for each
metropolitan area, one containing the travel times to jobs and the other
to groceries stores \footnote{The information from Toronto, Hamilton,
  and Waterloo are aggregated at the Greater Golden Horseshoe travel
  time matrices, making it 20 datasets in total.}. These information can
be turned into spatial data using the socio-economic and demographic
datasets.

The socio-economic and demographic data contain other information
disaggregated at the DA level for each CMA. These refer to total
population by age groups, number of dwellings, number of individuals
below the Low Income Measure, etc. The aggregated statistics refer to
population aggregates (for the selected CMA's) and transit revenue and
ridership data aggregated by regional and national scale.

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. The Table \citet{tbl-table_1}
below summarizes the travel time estimates. The travel time datasets
contain, in total, 97,784,850 origin-destination pairs (observations)
from population to employment locations for all the DA's in the sample
and 18,519,897 pairs from population to groceries stores. Considering
all areas combined, the mean travel time to jobs was 80 and 76 for
groceries stores.

\begingroup
\setlength\LTleft{0\linewidth}
\setlength\LTright{0\linewidth}\fontsize{7.5pt}{9.0pt}\selectfont
\begin{longtable}{@{\extracolsep{\fill}}clrrrrrr}
\toprule
Study Region Name & Destination & Observations & Mean & Sd & P25 & P50 & P75 \\ 
\midrule\addlinespace[2.5pt]
All regions & Employment & 97,784,850 & 78 & 25 & 59 & 80 & 99 \\ 
Toronto & Employment & 41,038,062 & 82 & 25 & 64 & 85 & 103 \\ 
Montréal & Employment & 30,091,411 & 76 & 25 & 57 & 77 & 97 \\ 
Vancouver & Employment & 12,254,478 & 76 & 25 & 58 & 77 & 96 \\ 
Calgary & Employment & 3,022,244 & 75 & 23 & 59 & 75 & 91 \\ 
Ottawa & Employment & 2,936,912 & 74 & 24 & 56 & 74 & 93 \\ 
Edmonton & Employment & 2,057,486 & 70 & 23 & 53 & 70 & 87 \\ 
Québec City & Employment & 1,473,702 & 72 & 25 & 54 & 72 & 93 \\ 
Winnipeg & Employment & 1,574,205 & 62 & 22 & 46 & 61 & 76 \\ 
Hamilton & Employment & 2,089,172 & 82 & 28 & 61 & 87 & 106 \\ 
Waterloo & Employment & 552,826 & 71 & 27 & 49 & 69 & 93 \\ 
London & Employment & 415,174 & 61 & 21 & 46 & 60 & 74 \\ 
Halifax & Employment & 279,178 & 66 & 25 & 47 & 65 & 84 \\ 
All regions & Groceries Stores & 18,519,897 & 76 & 26 & 56 & 76 & 97 \\ 
Toronto & Groceries Stores & 8,512,874 & 80 & 25 & 61 & 82 & 101 \\ 
Montréal & Groceries Stores & 2,993,965 & 73 & 26 & 53 & 74 & 94 \\ 
Vancouver & Groceries Stores & 4,540,106 & 73 & 25 & 53 & 74 & 93 \\ 
Calgary & Groceries Stores & 502,757 & 70 & 22 & 54 & 70 & 86 \\ 
Ottawa & Groceries Stores & 619,791 & 72 & 24 & 55 & 72 & 91 \\ 
Edmonton & Groceries Stores & 324,030 & 68 & 23 & 51 & 67 & 84 \\ 
Québec City & Groceries Stores & 234,600 & 70 & 26 & 51 & 70 & 89 \\ 
Winnipeg & Groceries Stores & 362,566 & 58 & 21 & 43 & 57 & 72 \\ 
Hamilton & Groceries Stores & 281,118 & 84 & 29 & 60 & 91 & 110 \\ 
Waterloo & Groceries Stores & 44,726 & 66 & 28 & 44 & 62 & 90 \\ 
London & Groceries Stores & 52,617 & 55 & 19 & 42 & 55 & 67 \\ 
Halifax & Groceries Stores & 50,747 & 63 & 26 & 42 & 63 & 83 \\ 
\bottomrule
\end{longtable}
\endgroup

>>>>

The objective of this paper is to describe the \{canaccesR\} open data
package. Its main contents are a set of public transit travel time
matrices (TTM) estimates to employment and groceries stores from the 12
largest Canadian metropolitan areas in 2019 and 2023, representing
approximately 55\% of the Canadian population. The results are provided
at the Dissemination Areas (DA) \footnote{Dissemination Areas are the
  smallest publicly available spatial unit provided by Statistics Canada
  \citep{governmentofcanadaDictionaryCensusPopulation2021a}.} level,
yielding origin-destination pairs containing associated travel time by
public transit, population, total employment, mode share, and other
relevant census variables and spatial shape boundaries for each
metropolitan area. This data package was created by leveraging expertise
in data science, R programming, and transportation analysis. It includes
.Rmd notebooks written for computing travel time matrices for large sets
of origin-destination pairs using the \{r5r\} R package
\citep{pereiraR5rRapidRealistic2021}. Overall, \{canaccesR\} offers 52
complementary objects ready for temporal and spatial analysis. The
package is an analysis-read product based on a fusion of data sources,
including public transit schedules (GTFS files), road and transit
networks, census data, and filtered business location data.

TTMs are a core piece of information required for estimating spatial
interaction. Accessibility -the \emph{potential} spatial interaction
offered by the transportation system to reach destinations
\citep{paezMeasuringAccessibilityPositive2012}- is an example of a
measure that requires TTMs. Recent efforts following an open-source and
transparent philosophy have been made to disseminate useful data and
information on transportation in the Canadian context
\citep{soukhovTTS2016RDataSet2023}. However, despite these initiatives,
pre-processed and available data that allow for the ease estimation of
accessibility indicators are still scarce. Within this context, we
expect to help filling this gap by the processing raw into user-ready
data and making them publicly available to advance knowledge on the
field. Our main contribution is to provide analysis-ready data for
Canada's largest cities on the topic of transportation accessibility,
thus making urban analytics in the country more accessible and
contributing to future research and data-based decision making.

@@

The locations included in the data package comprise the 12 largest
(population-wise) Census metropolitan areas (CMA's) based on the 2016
Canadian Census \citep{governmentofcanada2016CensusPopulation2016}
\footnote{We included Oshawa as part of the Greater Toronto Area (GTA)
  because of its proximity. Similarly, we included Abbotsford-Mission as
  part of the Vancouver metropolitan area due to its proximity to a
  transit station on the region's West Coast Express commuter rail line.}.These
locations are the surrunding Toronto, Montreal, Vancouver,
Ottawa-Gatineau, Calgary, Edmonton, Quebec City, Winnipeg, Hamilton,
Kitchener-Cambridge-Waterloo, London, and Halifax areas. We used four
main data sources to construct the \{canaccesR\} data package: General
Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI's Enhanced
Points of Interest, and Statistics Canada Census data.

@@

Using the \{r5r\} R package, we estimated public transit travel times
for two destination types, grocery stores and jobs. For each amenity
type, we chose a likely travel time and day of the week. We set a 15
minutes time window and the maximum trip duration to 120 minutes. The
estimated times are the median of the 15 minute time window. For
groceries stores, we set the departure date to a weekend afternoon and
the departure time to between 12:00 PM to 12:15 PM on April 20, 2019 and
April 22, 2023. For employment, we ran the analysis on a typical weekday
morning rush-hour commute, more specifically 8:00 to 8:15 AM departure
on Tuesday, April 16, 2019 and Tuesday, April 18, 2023 \footnote{The one
  exception is Quebec City, where the routing for 2019 occurs on a
  Saturday and Tuesday in June (instead of April) due to the GTFS data
  unavailability.}. In both cases, we assumed that walking was the mode
of travel from origin to transit stop and from transit stop to
destination. We aggregated all the resulting travel time matrices at the
Dissemination Area (DA) level, which comprise the fundamental unit of
analysis in data package.

@@

\begin{verbatim}
 [1] "census_data_cma_cal"   "census_data_cma_edm"   "census_data_cma_ggh"  
 [4] "census_data_cma_hal"   "census_data_cma_ldn"   "census_data_cma_mtl"  
 [7] "census_data_cma_ott"   "census_data_cma_que"   "census_data_cma_van"  
[10] "census_data_cma_win"   "census_data_da_cal"    "census_data_da_edm"   
[13] "census_data_da_ggh"    "census_data_da_hal"    "census_data_da_ldn"   
[16] "census_data_da_mtl"    "census_data_da_ott"    "census_data_da_que"   
[19] "census_data_da_van"    "census_data_da_win"    "population_statistics"
[22] "region_background_cal" "region_background_edm" "region_background_ggh"
[25] "region_background_hal" "region_background_ldn" "region_background_mtl"
[28] "region_background_ott" "region_background_que" "region_background_van"
[31] "region_background_win" "transit_statistics"    "travel_matrix_emp_cal"
[34] "travel_matrix_emp_edm" "travel_matrix_emp_ggh" "travel_matrix_emp_hal"
[37] "travel_matrix_emp_ldn" "travel_matrix_emp_mtl" "travel_matrix_emp_ott"
[40] "travel_matrix_emp_que" "travel_matrix_emp_van" "travel_matrix_emp_win"
[43] "travel_matrix_grc_cal" "travel_matrix_grc_edm" "travel_matrix_grc_ggh"
[46] "travel_matrix_grc_hal" "travel_matrix_grc_ldn" "travel_matrix_grc_mtl"
[49] "travel_matrix_grc_ott" "travel_matrix_grc_que" "travel_matrix_grc_van"
[52] "travel_matrix_grc_win"
\end{verbatim}

\begin{verbatim}
 [1] 11 12 13 14 15 16 17 18 19 20
\end{verbatim}

Specifically, the package contains the following contents: 10 data.frame
objects containing the calculated public travel times from DA centroids
to grocery stores and another 10 for DA centroids to DA centroids.
Notably, the \{travel\_matrix\_grc\_ggh\} data.frame, is named after the
acronym for the Greater Golden Horseshoe area, which includes the CMA
regions of Toronto, Hamilton, and Kitchener-Cambridge-Waterloo. Hence,
the 20 travel time data.frames represent public transit travel times for
all 12 CMAs across two sets of destinations. Next, 10 sf objects
represent census data for each DA, including dwelling counts, population
by age bracket, single-parent-headed households, low-income prevalence,
official language knowledge, housing- quality, ownership and
affordability variables, visible minority, newcomer- and immigration-
related variables, educational attainment, and commuting mode shares.
Furthermore, 10 and 10 sf objects represent the CMA areas' boundaries
and backgrounds for plotting the data spatially, respectively. Finally,
2 data.frames contain aggregated population and transit statistics at
the CMA level.

--\textgreater{}

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. Table \citet{tbl-table_1}
summarizes the travel time estimates. The travel time objects contain,
in total, 97,784,850 origin-destination pairs (observations) from
population to employment locations for all the DA's in the sample and
18,519,897 pairs from population to groceries stores. Considering all
areas combined, the mean travel time to jobs was 80 and 76 for groceries
stores.

\begin{table}[!t]
\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}clrrrrrr}
\toprule
Study Region Name & Destination & Observations & Mean & Sd & P25 & P50 & P75 \\ 
\midrule\addlinespace[2.5pt]
All regions & Employment & 97,784,850 & 78 & 25 & 59 & 80 & 99 \\ 
Toronto & Employment & 41,038,062 & 82 & 25 & 64 & 85 & 103 \\ 
Montréal & Employment & 30,091,411 & 76 & 25 & 57 & 77 & 97 \\ 
Vancouver & Employment & 12,254,478 & 76 & 25 & 58 & 77 & 96 \\ 
Calgary & Employment & 3,022,244 & 75 & 23 & 59 & 75 & 91 \\ 
Ottawa & Employment & 2,936,912 & 74 & 24 & 56 & 74 & 93 \\ 
Edmonton & Employment & 2,057,486 & 70 & 23 & 53 & 70 & 87 \\ 
Québec City & Employment & 1,473,702 & 72 & 25 & 54 & 72 & 93 \\ 
Winnipeg & Employment & 1,574,205 & 62 & 22 & 46 & 61 & 76 \\ 
Hamilton & Employment & 2,089,172 & 82 & 28 & 61 & 87 & 106 \\ 
Waterloo & Employment & 552,826 & 71 & 27 & 49 & 69 & 93 \\ 
London & Employment & 415,174 & 61 & 21 & 46 & 60 & 74 \\ 
Halifax & Employment & 279,178 & 66 & 25 & 47 & 65 & 84 \\ 
All regions & Groceries Stores & 18,519,897 & 76 & 26 & 56 & 76 & 97 \\ 
Toronto & Groceries Stores & 8,512,874 & 80 & 25 & 61 & 82 & 101 \\ 
Montréal & Groceries Stores & 2,993,965 & 73 & 26 & 53 & 74 & 94 \\ 
Vancouver & Groceries Stores & 4,540,106 & 73 & 25 & 53 & 74 & 93 \\ 
Calgary & Groceries Stores & 502,757 & 70 & 22 & 54 & 70 & 86 \\ 
Ottawa & Groceries Stores & 619,791 & 72 & 24 & 55 & 72 & 91 \\ 
Edmonton & Groceries Stores & 324,030 & 68 & 23 & 51 & 67 & 84 \\ 
Québec City & Groceries Stores & 234,600 & 70 & 26 & 51 & 70 & 89 \\ 
Winnipeg & Groceries Stores & 362,566 & 58 & 21 & 43 & 57 & 72 \\ 
Hamilton & Groceries Stores & 281,118 & 84 & 29 & 60 & 91 & 110 \\ 
Waterloo & Groceries Stores & 44,726 & 66 & 28 & 44 & 62 & 90 \\ 
London & Groceries Stores & 52,617 & 55 & 19 & 42 & 55 & 67 \\ 
Halifax & Groceries Stores & 50,747 & 63 & 26 & 42 & 63 & 83 \\ 
\bottomrule
\end{tabular*}
\end{table}
